Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-05-05 document/ms.tex
refs

The transit method of exoplanet detection and characterization has been
demonstrated as a powerful method for building systematic catalogs of
exoplanets.
Despite the great success of the \kepler\ Mission with thousands of planet
discoveries \citep{Burke:2014}, current methods for exoplanet discoveries are
currently limited in the range of orbital periods that can be studied.
Specifically, the standard transit search procedures only discover signals
with at least three observed transits \citep[for example][]{Petigura:2013,
Burke:2014}.
For \kepler, with a baseline of about four years, this sets an upper limit on
the detectable periods of just over a year.
In the Solar System, Jupiter---with a period of 12 years---dominates the
planetary dynamics and, since it would only exhibit at most one transit in the
\kepler\ data, it would be missed by most existing transit search procedures.
It is possible to discover long-period planets like this using targeted radial
velocity (RV) surveys \citep[for example][]{Butler:2006, Knutson:2014} but the
cost of implementing a systematic RV search is substantially higher than
searching the existing and forthcoming photometric data for single transits.

@@

Nearly every transit search algorithm is built on the same principles and many
of the same decisions are made.
In particular, at the heart of most methods is a matched filtering step where
the likelihood of an approximate transit model is computed on a grid in the
physical parameters \citep[\kepler\ Data Processing
Handbook\footnote{\url{https://archive.stsci.edu/kepler/manuals/KSCI-19081-001_Data_Processing_Handbook.pdf}};][]{%
Petigura:2013, Huang:2013, Dressing:2015, Foreman-Mackey:2015}.
Using these methods, and substantial hand-curation, some long-period
transiting candidates have been published \citep[for example][]{Batalha:2013,
Huang:2013, Kipping:2014a}.
For more recent data releases, the community has settled on more conservative
selection criteria where candidates are required to have multiple transits
\citep[for example][]{Petigura:2013, Burke:2014}.
Even the \project{QATS} algorithm \citep{Carter:2013} for finding
quasiperiodic transits builds on much the same infrastructure.
This means that there has never been a systematic search for single transits
in the full \kepler\ dataset.

@@

We estimate the precision of each RF classification model on the validation
set as described in \sect{train}.
As discussed in that \sectionname, we choose a target precision to
100~percent and set the score threshold accordingly.
This estimate is only applicable to the test set under two major assumptions.
The first is that all transiting planets with more than one transit and a
large signal-to-noise ratio have been previously discovered by one of the many
transit searches applied the \kepler\ dataset \citep{Burke:2014}.
This is a safe assumption because, for large planets, these surveys are all
largely complete out to periods yielding only 2 or 3 transits in the \kepler\
baseline \citep[for example][]{Petigura:2013}.
The second assumption is not always valid so it makes this estimate of the
precision only approximate and not completely reliable.
This assumption is that the noise processes in the data are stationary.
In other words, the transit-like noise signals in the training and validation
sets must be ``similar'' to the signals in the test section.
This constraint is nearly satisfied when the PDC light curves are used
(\sect{data}) and the data are split carefully (\sect{split}).
As demonstrated by the mis-classifications shown in \sect{demo}, however, it
is clear that violations to this rule do exist in the data and final vetting
of the candidates must be applied using a physical understanding of the
signals.

>>>>

The transit method of exoplanet detection and characterization has been
demonstrated as a powerful method for building systematic catalogs of
exoplanets.
Despite the great success of the \kepler\ Mission with thousands of planet
discoveries \citep{Burke:2014, Rowe:2015}, current methods for exoplanet
discoveries are currently limited in the range of orbital periods that can be
studied.
Specifically, the standard transit search procedures only discover signals
with at least three observed transits \citep[for example][]{Petigura:2013,
Burke:2014, Rowe:2015}.
For \kepler, with a baseline of about four years, this sets an upper limit on
the detectable periods of just over a year.
In the Solar System, Jupiter---with a period of 12 years---dominates the
planetary dynamics and, since it would only exhibit at most one transit in the
\kepler\ data, it would be missed by most existing transit search procedures.
It is possible to discover long-period planets like this using targeted radial
velocity (RV) surveys \citep[for example][]{Butler:2006, Knutson:2014} but the
cost of implementing a systematic RV search is substantially higher than
searching the existing and forthcoming photometric data for single transits.

@@

Nearly every transit search algorithm is built on the same principles and many
of the same decisions are made.
In particular, at the heart of most methods is a matched filtering step where
the likelihood of an approximate transit model is computed on a grid in the
physical parameters \citep[\kepler\ Data Processing
Handbook\footnote{\url{https://archive.stsci.edu/kepler/manuals/KSCI-19081-001_Data_Processing_Handbook.pdf}};][]{%
Petigura:2013, Huang:2013, Dressing:2015, Foreman-Mackey:2015}.
Using these methods, and substantial hand-curation, some long-period
transiting candidates have been published \citep[for example][]{Batalha:2013,
Huang:2013, Kipping:2014a}.
For more recent data releases, the community has settled on more conservative
selection criteria where candidates are required to have multiple transits
\citep[for example][]{Petigura:2013, Burke:2014, Rowe:2015}.
Even the \project{QATS} algorithm \citep{Carter:2013} for finding
quasiperiodic transits builds on much the same infrastructure.
This means that there has never been a systematic search for single transits
in the full \kepler\ dataset.

@@

We estimate the precision of each RF classification model on the validation
set as described in \sect{train}.
As discussed in that \sectionname, we choose a target precision to
100~percent and set the score threshold accordingly.
This estimate is only applicable to the test set under two major assumptions.
The first is that all transiting planets with more than one transit and a
large signal-to-noise ratio have been previously discovered by one of the many
transit searches applied the \kepler\ dataset \citep{Burke:2014, Rowe:2015}.
This is a safe assumption because, for large planets, these surveys are all
largely complete out to periods yielding only 2 or 3 transits in the \kepler\
baseline \citep[for example][]{Petigura:2013}.
The second assumption is not always valid so it makes this estimate of the
precision only approximate and not completely reliable.
This assumption is that the noise processes in the data are stationary.
In other words, the transit-like noise signals in the training and validation
sets must be ``similar'' to the signals in the test section.
This constraint is nearly satisfied when the PDC light curves are used
(\sect{data}) and the data are split carefully (\sect{split}).
As demonstrated by the mis-classifications shown in \sect{demo}, however, it
is clear that violations to this rule do exist in the data and final vetting
of the candidates must be applied using a physical understanding of the
signals.
